% last updated in April 2002 by Antje Endemann
% Based on CVPR 07 and LNCS, with modifications by DAF, AZ and elle, 2008 and AA, 2010, and CC, 2011; TT, 2014; AAS, 2016

\documentclass[runningheads]{llncs}
\usepackage{graphicx}
\usepackage{amsmath,amssymb} % define this before the line numbering.
\usepackage{ruler}
\usepackage{color}
\usepackage{wrapfig}
\usepackage{makecell}
\usepackage{rotating}
\usepackage{enumitem}
\usepackage{multirow}
\usepackage{booktabs,tabu}
\usepackage[width=122mm,left=12mm,paperwidth=146mm,height=193mm,top=12mm,paperheight=217mm]{geometry}

\usepackage{cite} % automatically re-orders citation numbers

\usepackage{floatrow}

\newcommand\todo[1]{\textcolor{red}{#1}}

\newcommand\degr[0]{^{\circ}}


\begin{document}
% \renewcommand\thelinenumber{\color[rgb]{0.2,0.5,0.8}\normalfont\sffamily\scriptsize\arabic{linenumber}\color[rgb]{0,0,0}}
% \renewcommand\makeLineNumber {\hss\thelinenumber\ \hspace{6mm} \rlap{\hskip\textwidth\ \hspace{6.5mm}\thelinenumber}}
% \linenumbers
\pagestyle{headings}
\mainmatter

\def\ECCV18SubNumber{1114}

\title{Deep Photometric Stereo on a Sunny Day}
% Important keywords to highlight somehow in the title: Outdoor, One-Day, Sunny, PS, Deep

\titlerunning{ECCV-18 submission ID \ECCV18SubNumber}
\authorrunning{ECCV-18 submission ID \ECCV18SubNumber}

\author{Anonymous ECCV submission}
\institute{Paper ID \ECCV18SubNumber}


\maketitle

%  Given multiple images with different illumination taken from a single viewpoint, it is possible to recover the normals of the surfaces present in the scene. 
% Its application to outdoor scenes has seen recent interest. Unfortunately, most successful outdoor PS techniques requires either months of data or specific events to happen like an alternance of cloud occluding the sun.
\begin{abstract}
Photometric Stereo in outdoor illumination remains a challenging, ill-posed problem. Indeed, it has been shown that the scene structure cannot be recovered unambiguously from the photometric cues alone when it is lit by the sun over the course of a single day. In this paper, we present a CNN-based technique for photometric stereo on a single sunny day. We solve the ambiguity issue by combining photometric cues with prior knowledge on material properties, local surface geometry and the natural variations in outdoor lighting. To train the CNN, we create a dataset of realistic synthetic renders with both diffuse and specular materials. Given eight outdoor images taken during a single sunny day, our method robustly estimates the scene surface normals. Our approach does not require precise geolocation to work and significantly outperforms several state-of-the-art methods on images with real lighting. This shows that our CNN can combine efficiently learned priors and photometric cues available during a single sunny day.

\keywords{photometric stereo, deep learning, outdoor lighting}
\end{abstract}

% Even though the technique is robust to glossy materials, it requires a scene with uniform albedo and a camera pointing toward the North. 


\input{intro}
\input{relatedwork}
\input{method}
\input{results}
\input{analysis}
\input{conclusion}


\bibliographystyle{splncs}
\bibliography{bibliography}
\end{document}
